---
title: "Project3"
author: "Gabrielle Bartomeo, Binish Chandry,Zach Dravis,Burcu Kaniskan,Niteen Kumar, Betsy Rosalen"
date: "3/21/2018"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_section:  true
    theme: cerulean
    highlight:  tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries

```{r message=FALSE}
library(knitr)
library(tidyverse)
library(DT)
library(kableExtra)
library(lubridate)
```

#Importing data
The data set used in this project is [Kaggle ML and Data Science Survey 2017](https://www.kaggle.com/kaggle/kaggle-survey-2017/feed). The survey was stored in 2 different data sets:a) multiple choice items, b) free-response items. Kaggle stored each data in csv format. We dowloaded the multiple choice item survey results in csv format and placed it in our [GitHub repo](https://github.com/betsyrosalen/DATA_607_Project_3)

Importing Multiple Choice data
```{r message=FALSE ,warning=FALSE}
linkMC<-"https://raw.githubusercontent.com/betsyrosalen/DATA_607_Project_3/master/project3_master/rawdata/multipleChoiceResponses.csv"
#importing MC items
MC<-read_csv (linkMC)
dim(MC)
#lets create a unique ID variable 
MC$id <- seq.int(nrow(MC))
```

Ignore this codeImporting conversionrates data incase we want to do analyses 
```{r message=FALSE ,warning=FALSE}
# link_conversion<-"https://raw.githubusercontent.com/betsyrosalen/DATA_607_Project_3/master/project3_master/rawdata/conversionRates.csv"
# #importing MC items
# conversion<-read_csv (link_conversion)
# dim(conversion)
# #lets create a unique ID variable 
# conversion$id <- seq.int(nrow(conversion))
```



#Subsetting the data

We subset our data for the columns we need

```{r}
df<-MC %>%
select (WorkToolsSelect, id, Country, JobFunctionSelect, CurrentJobTitleSelect, LanguageRecommendationSelect, LearningPlatformSelect, FormalEducation, LearningPlatformSelect:LearningPlatformUsefulnessYouTube, CompensationAmount, CompensationCurrency, EmploymentStatus, TitleFit, CodeWriter, LearningDataScience, CurrentEmployerType, StudentStatus)
colnames(df)
```



# Research Questions
This project will answer this globalresearch question **Which are the most values data science skills?**
The following 6 research questions will provide answer to this global question.

##Research Question1

What is the relationship between the most popular platforms for learning DS and X (Niteen)?
Alternatively phrased: What data science learning resources and which locations of open data are utilized by people of varying levels of education? (delete me if you need to!)

## Variables and their definition 

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```


## Research Question2
Does survey takers' formal education has any relationship to the ML/DS method he or she is most excited about learning in the next year? (Binish)

##Variables and their definition

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```

## Research Question3
What are the most frequently used DS methods? Where is the most time spent in terms of working with data? Do either of these correlate with job title or level of education? (Zach)

## Variables and their definition 

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```


## Research Question4
Is there a difference between what 'Learners' think are the important skills to learn and what employed Data Scientists say are the skils and tools they are using? (Betsy)

## Variables and their definition

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```

## Research Question5
Is there any interaction between the Kaggle survey takers' program language use (R or Python) and their recommended program languages? (e.g. R users recommending R more than Python users recommending Python) (Burcu)

## Variables and their definition 

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data


```{r}

#creating a new variable called work_tools where the original column values are split
RQ5<-df %>% 
  mutate(work_tools = strsplit(as.character(WorkToolsSelect), ",")) %>% 
  unnest(work_tools)

#checking the top 100 rows for the newly created work_tools variable using top_n function
RQ5 %>%
select (work_tools,WorkToolsSelect, id)%>%
filter (id %in% c(1:5))%>%
 datatable()

```
### Exploratory Data Analysis (EDA)

## Research Question6
Of those receiving pay in US Dollars, is Python or R overall most profitable for a Kaggle survey taker? (Gabby)

## Variables and their definition

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

```{r}
RQ6 <- filter(RQ5, !is.na(WorkToolsSelect)) # Filters out all columns with NA in the WorkToolsSelect column
RQ6 <- RQ6 %>%
  filter(CompensationCurrency == "USD") %>% # Makes sure to only use rows whose currency is in USD
  filter(work_tools == "Python" | work_tools == "R") %>% # The work tools are R or Python, period.
  select(id, work_tools, CompensationAmount) # Only have three rows to work with
RQ6_ids <- select(filter(as.data.frame(table(RQ6$id)), Freq == 1), Var1) # Only want people who use R or Python EXCLUSIVELY, not R and/or Python
RQ6_ids <- droplevels(RQ6_ids)$Var1 # Removed the levels so we can actually get the IDs
RQ6 <- filter(RQ6, id %in% RQ6_ids) # Only keep those rows whose id are inside of list of ids with R or Python exclusively used at work
RQ6 <- select(RQ6, -id) # No use for the ID anymore, it's done its job
RQ6$CompensationAmount <- gsub(",", "", RQ6$CompensationAmount) # Removed the commas from the compensation amount to prep for numeric transformation
RQ6$CompensationAmount <- as.numeric(RQ6$CompensationAmount) # made the column into a numeric for easier mathematical comparison and sorting
RQ6 <- filter(RQ6, CompensationAmount < 9999999) # ... let's just be a little realistic, nobody is earning more than fifteen million a year at this point in time or prior to it, and this one-dollar-off-from-a-million entry is an anomaly in the data set
rm(RQ6_ids) # remove the now-unused variable to save memory
```

### Exploratory Data Analysis (EDA)

```{r}
RQ6_boxplot <- ggplot(RQ6) +
  geom_boxplot( aes(x = factor(work_tools), 
                    y = CompensationAmount,
                    fill = factor(work_tools)
                     )
                  ) +
  scale_y_continuous(breaks=seq(0,2000000,25000)) +
  labs( x = "Programming Language",
        y = "Annual Compensation in USD",
        fill = "Programming Language")
RQ6_boxplot_ylim <- boxplot.stats(RQ6$CompensationAmount)$stats[c(1, 5)]
RQ6_boxplot <- RQ6_boxplot + coord_cartesian(ylim = RQ6_boxplot_ylim*1.05)
RQ6_boxplot
```

The average survey taker who used Python in their job made approximately \$`r formatC(colMeans(select(filter(RQ6, work_tools=="Python"),CompensationAmount))-colMeans(select(filter(RQ6, work_tools=="R"),CompensationAmount)), digits=2, format="f", big.mark=",")` more than the average survey taker who used R in their job. While R users overall had a higher base pay - to the tune of $`r formatC(quantile(select(filter(RQ6, work_tools=="R"),CompensationAmount)[[1]])["25%"]-quantile(select(filter(RQ6, work_tools=="Python"),CompensationAmount)[[1]])["25%"], digits=2, format="f", big.mark=",")` more than their Python counterparts - their ability to achieve growth in salary was noticeably stymied in comparison. Outliers aside, if the data collected is to be considered representative of the data science population, there is indication that a prospective Data Scientist should learn R first for a higher initial salary, and then learn Python to increase their chance of obtaining a job with more growth potential.
