---
title: "Project3"
author: "Gabrielle Bartomeo, Binish Chandry,Zach Dravis,Burcu Kaniskan,Niteen Kumar, Betsy Rosalen"
date: "3/21/2018"
output: 
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    number_section:  true
    theme: cerulean
    highlight:  tango
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries

```{r message=FALSE}
library(knitr)
library(tidyverse)
library(DT)
library(kableExtra)
library(lubridate)
```

#Importing data
The data set used in this project is [Kaggle ML and Data Science Survey 2017](https://www.kaggle.com/kaggle/kaggle-survey-2017/feed). The survey was stored in 2 different data sets:a) multiple choice items, b) free-response items. Kaggle stored each data in csv format. We dowloaded the multiple choice item survey results in csv format and placed it in our [GitHub repo](https://github.com/betsyrosalen/DATA_607_Project_3)

Importing Multiple Choice data
```{r message=FALSE ,warning=FALSE}
linkMC<-"https://raw.githubusercontent.com/betsyrosalen/DATA_607_Project_3/master/project3_master/rawdata/multipleChoiceResponses.csv"
#importing MC items
MC<-read_csv (linkMC)
dim(MC)
#lets create a unique ID variable 
MC$id <- seq.int(nrow(MC))
```

Ignore this codeImporting conversionrates data incase we want to do analyses 
```{r message=FALSE ,warning=FALSE}
# link_conversion<-"https://raw.githubusercontent.com/betsyrosalen/DATA_607_Project_3/master/project3_master/rawdata/conversionRates.csv"
# #importing MC items
# conversion<-read_csv (link_conversion)
# dim(conversion)
# #lets create a unique ID variable 
# conversion$id <- seq.int(nrow(conversion))
```



#Subsetting the data

We subset our data for the columns we need

```{r}
df<-MC %>%
select (WorkToolsSelect, id, Country, JobFunctionSelect, CurrentJobTitleSelect, LanguageRecommendationSelect, LearningPlatformSelect, FormalEducation, LearningPlatformSelect:LearningPlatformUsefulnessYouTube, CompensationAmount, CompensationCurrency)
colnames(df)
```



# Research Questions
This project will answer this globalresearch question **Which are the most values data science skills?**
The following 6 research questions will provide answer to this global question.

##Research Question1

What is the relationship between the most popular platforms for learning DS and X (Niteen)?

## Variables and their definition 

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```


## Research Question2
Does survey takers' formal education has any relationship to the ML/DS method he or she is most excited about learning in the next year? (Binish)

##Variables and their definition

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```

## Research Question3
What are the most frequently used DS methods? Where is the most time spent in terms of working with data? Do either of these correlate with job title or level of education? (Zach)

## Variables and their definition 

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```


## Research Question4
Is there a difference between what 'Learners' think are the important skills to learn and what employed Data Scientists say are the skils and tools they are using? (Betsy)

## Variables and their definition

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```

## Research Question5
Is there any interaction between the Kaggle survey takers' program language use (R or Python) and their recommended program languages? (e.g. R users recommending R more than Python users recommending Python) (Burcu)

## Variables and their definition 

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data


```{r}

#creating a new variable called work_tools where the original column values are split
RQ5<-df %>% 
  mutate(work_tools = strsplit(as.character(WorkToolsSelect), ",")) %>% 
  unnest(work_tools)

#checking the top 100 rows for the newly created work_tools variable using top_n function
RQ5 %>%
select (work_tools,WorkToolsSelect, id)%>%
filter (id %in% c(1:5))%>%
 datatable()

```
### Exploratory Data Analysis (EDA)

## Research Question6
Of those receiving pay in US Dollars, is Python or R overall most profitable for a Kaggle survey taker? (Gabby)

## Variables and their definition

This section will describe the name of the variables and their labels (as reported in schema doc) and how the values were codes (etc yes, no, select all)

### Manipulating data

### Exploratory Data Analysis (EDA)

```{r}

```
